### 2022-07-19 ETI/cover_letter.tex
TOC figure

\documentclass[12pt]{dinbrief}

%\usepackage{english}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{color}

\usepackage[unicode=true, pdfusetitle, bookmarks=true, bookmarksnumbered=false, bookmarksopen=false, breaklinks=false, pdfborder={0 0 0}, pdfborderstyle={}, backref=false, colorlinks=true] {hyperref}

%\usepackage[pdfborder={0 0 0}]{hyperref}

\date{\today}

\place{Prague}

\nowindowtics
\nowindowrules

\newcommand{\theeditor}{Professor Ravi Naidu}

\newcommand{\thejournal}{\textit{Environmental Technology \& Innovation}}
\newcommand{\cl}{\mathrm{Cl^-}}
\newcommand{\na}{\mathrm{Na^+}}

\newcommand{\ie}{\textit{i.~e.} } 
\enabledraftstandard

\begin{document}

\begin{letter}{
Dr. Oleg Rud\\
Department of Physical and Macromolecular Chemistry\\
Faculty of Science, Charles University\\
Hlavova 8, 128 43 Prague, Czech Republic\\
E-mail: oleg.rud@natur.cuni.cz\\
}
\vspace*{-1.5cm}

\opening{Dear \theeditor,}

Herewith we would like to submit our manuscript: \textit{Water desalination using polyelectrolyte hydrogel. Gibbs ensemble modeling}
authored by
Mikhail Laktionov, Lucie Nová and Oleg V. Rud,
 to be considered for publication in \thejournal. 
In the submitted manuscript we modeled hydrogel in Gibbs ensemble, \ie in equilibrium with a small amount of aqueous solution, and demonstrated that the compression of the gel in such an ensemble slightly decreases the salinity of the outside solution. 
We employed this phenomenon and modeled the process of water desalination as a cascade of hydrogel swellings and compressions changing the outside salinity from $\sim 0.1$ to 0.01 mol/l.

We believe that our work brings new insights into the complex problem of seawater desalination
that fits within the scope of \thejournal, and that it will be of interest to its broad readership.

\vspace{5ex}

Yours sincerely,\\
on behalf of all co-authors,

\hspace{80ex}\ O.V. Rud\\

%The advantage of the proposed desalination procedure over the forward osmosis is that the gel acts both as a draw solute and as a separation membrane. 
%That helps to avoid various blocking phenomena (characteristic of membrane processes), which required a regular chemical and mechanical cleaning.
%Moreover, the relatively low pressures needed for hydrogel compression allow to  maintain the hydrogel mechanical properties and integrity for its reuse in a serial cycle.

\end{letter}
\end{document}